---
title: "Generate laplacian eigenmaps embedding for GPLVM"
author: "Kieran Campbell"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_align: center
    toc: yes
  pdf_document:
    toc: yes
---

In this vignette we apply `embeddr` to the [Quake et al.](http://www.nature.com/nature/journal/v509/n7500/full/nature13173.html) dataset. For copyright reasons we do not include the raw data in the package. It can be downloaded from [here](http://www.nature.com/nature/journal/v509/n7500/extref/nature13173-s4.txt) and converted to an `SCESet` using [the accompanying vignette](https://github.com/kieranrcampbell/embeddr/blob/master/vignettes/quake_to_scater.Rmd). 

In this example we have saved it to `~/oxford/embeddr/data/lsce.Rdata`, though this should be changed to where you save it.

```{r load-lib, message=FALSE, warning=FALSE}
library(plyr)
library(dplyr)
library(devtools)
library(data.table)
library(ggplot2)
library(reshape2)
library(splines)
library(corrplot)
library(scater)
library(gplots)
library(NMF)
library(monocle)
library(embeddr)
library(rhdf5)

set.seed(123)
```

```{r load-data, cache=TRUE}
## load object 'lsce' 
#data_path <- '~/oxford/embeddr/data/lsce.Rdata'
data_path <- '/net/isi-scratch/kieran/embeddr/data/lsce.Rdata'
load(data_path)
lsce@lowerDetectionLimit <- 0 # from paper
```

## Building the embedding


Next we want to fit a noise model to find highly variable genes:
```{r noise-model, cache=TRUE, fig.align='center'}
Lmeans <- rowMeans(exprs(lsce))
Lvars <- apply(exprs(lsce), 1, var)

to_use <- Lmeans > 0.5 & Lvars > 0 

LCV <- Lvars[to_use] / Lmeans[to_use]^2

df_var <- data.frame(CV2 = LCV, m=Lmeans[to_use])

fit_loglin <- nls(CV2 ~ a + b / m, data = df_var, start=c(a=80, b=5)) 

f <- function(x) coef(fit_loglin)[1] + coef(fit_loglin)[2] / x

is_het <- LCV > 1.5 * predict(fit_loglin)
df_var$is_het <- is_het

ggplot(df_var, aes(x=m, y=CV2, color=is_het)) + geom_point() +
  stat_function(fun=f, color='black') + theme_bw()

het_genes <- names(which(is_het))
het_indices <- match(het_genes, rownames(lsce))
```

Now we can apply laplacian eigenmaps using a nearest-neighbour graph:

```{r le1, cache=TRUE, fig.align='center', message=FALSE, warning=FALSE}
lsce <- embeddr(lsce, genes_for_embedding = het_indices, nn = 10)

phenoData(lsce)$cluster <- pData(lsce)$putative_cell_type
plot_embedding(lsce, color_by = 'putative_cell_type')
plot_graph(lsce)
```


It is obvious we have two very separate trajectories; let's consider the pseudotemporal ordering of one only:

```{r cluster-plot, cache=TRUE, fig.align='center', message=FALSE}
set.seed(123)
lsce <- cluster_embedding(lsce, method='mm')#, k = 4)
plot_embedding(lsce)

cl1 <- pData(lsce)$cluster %in% 1:3
lsce_13 <- lsce[,cl1]

lsce_13 <- fit_pseudotime(lsce_13)
lsce_13_tmp <- lsce_13
phenoData(lsce_13_tmp)$cluster <- plyr::mapvalues(pData(lsce_13_tmp)$cluster,
                                                 from=1:3,
                                                 to=c('AT1','BP','AT2'))

plot_embedding(lsce_13_tmp)
plot_pseudotime_density(lsce_13_tmp, color_by = 'cluster')
```


```{r save-embeddings, cache=TRUE, warning=FALSE, message=FALSE}
h5file <- "/net/isi-scratch/kieran/GP/gpseudotime/data/embeddings.h5"
h5createFile(h5file)
h5createGroup(h5file, "lung")
h5write(redDim(lsce_13_tmp), h5file, "lung/embedding")
h5write(pseudotime(lsce_13_tmp), h5file, "lung/pseudotime")
```
  
